Automated segmentation and description of the internal morphology of human permanent teeth by means of micro-CT

### 2020-11-10 manuscript.tex
Merge branch 'rootstock-2020-11-10' [ci skip]
This build is based on
https://github.com/habi/zmk-tooth-cohort-method-manuscript/commit/877b421bd54e45c8867c3532d8125a141e25bf7d.

This commit was created by the following CI build and job:
https://github.com/habi/zmk-tooth-cohort-method-manuscript/commit/877b421bd54e45c8867c3532d8125a141e25bf7d/checks
https://github.com/habi/zmk-tooth-cohort-method-manuscript/runs/355900482

A preprint of this manuscript is available on \emph{bioRxiv} at \url{https://doi.org/10.1101/2020.10.27.356998} and has been submitted to \href{https://bmcoralhealth.biomedcentral.com/}{\emph{BMC Oral Health}}.
This version of the manuscript \href{https://github.com/habi/zmk-tooth-cohort-method-manuscript/compare/f9b3333c625228f4f28211a34f381b9f87c430b0...6cc6138d341a45345e2ef808848b1e1a2015baac}{has been updated} since the preprint deposition.

This manuscript
(\href{https://habi.github.io/zmk-tooth-cohort-method-manuscript/v/6cc6138d341a45345e2ef808848b1e1a2015baac/}{permalink})
was automatically generated
from \href{https://github.com/habi/zmk-tooth-cohort-method-manuscript/tree/6cc6138d341a45345e2ef808848b1e1a2015baac}{habi/zmk-tooth-cohort-method-manuscript@6cc6138}
on November 5, 2020.

@@

A total of 104 extracted human permanent mandibular canines were collected from university medical centers in southwest Germany and Switzerland.
All teeth included were extracted for reasons unrelated to the study and are so-called excess material making an institutional review board approval unnecessary for the purpose of this investigation.
The teeth were single-rooted and investigated according to their morphologic criteria.
Inclusion criteria for teeth selection were complete coronal and root development and the absence of root fracture and resorption, coronal and radicular caries and endodontic treatment.
Calculus as well as hard and soft tissue was removed using an ultrasonic scaler.
Afterwards, the teeth were placed for one hour in a 3\% hydrogen peroxide ultrasonic bath and then stored in 70\% ethanol {[}\protect\hyperlink{ref-1HkTk4VKM}{5}, \protect\hyperlink{ref-OttTicsv}{12}, \protect\hyperlink{ref-amSKhLb1}{15}, \protect\hyperlink{ref-ZTgDEakx}{17}, \protect\hyperlink{ref-1Gn1pVZXt}{19}{]}.

@@

\hypertarget{acknowledgements}{%
\subsection{Acknowledgements}\label{acknowledgements}}

We thank Oleksiy-Zakhar Khoma (Institute of Anatomy, University of Bern, Switzerland) for designing \href{https://github.com/TomoGraphics/Hol3Drs/blob/master/STL/tooth_holder3.stl}{the sample holder that we used for scanning the teeth}.
We thank Jennifer Fazzari (Institute of Anatomy, University of Bern, Switzerland) for proof-reading the manuscript.
We thank Andrea Anderegg and Michael Stiebritz (both Department of Restorative, Preventive and Pediatric Dentistry, School of Dental Medicine, University of Bern, Switzerland) and Valentin Djonov (Institute of Anatomy, University of Bern, Switzerland) for their kind support.
This research did not receive any specific grant from funding agencies in the public, commercial, or not-for-profit sectors.

\hypertarget{contributions}{%
\subsection{Contributions}\label{contributions}}

@@

\hypertarget{conflict-of-interest-statement}{%
\subsection{Conflict of interest statement}\label{conflict-of-interest-statement}}

>>>>

A preprint of this manuscript is available on \emph{bioRxiv} at \url{https://doi.org/10.1101/2020.10.27.356998} and has been submitted to \href{https://bmcoralhealth.biomedcentral.com/}{\emph{BMC Oral Health}}.
This version of the manuscript \href{https://github.com/habi/zmk-tooth-cohort-method-manuscript/compare/f9b3333c625228f4f28211a34f381b9f87c430b0...877b421bd54e45c8867c3532d8125a141e25bf7d}{has been updated} since the preprint deposition.

This manuscript
(\href{https://habi.github.io/zmk-tooth-cohort-method-manuscript/v/877b421bd54e45c8867c3532d8125a141e25bf7d/}{permalink})
was automatically generated
from \href{https://github.com/habi/zmk-tooth-cohort-method-manuscript/tree/877b421bd54e45c8867c3532d8125a141e25bf7d}{habi/zmk-tooth-cohort-method-manuscript@877b421}
on November 10, 2020.

@@

A total of 104 extracted human permanent mandibular canines were collected from university medical centers in southwest Germany and Switzerland.
All included teeth were extracted for reasons unrelated to the study and are so-called excess material making an institutional review board approval unnecessary for the purpose of this investigation.
The teeth were single-rooted and investigated according to their morphologic criteria.
Inclusion criteria for teeth selection were complete coronal and root development and the absence of root fracture and resorption, coronal and radicular caries and endodontic treatment.
Calculus as well as hard and soft tissue was removed using an ultrasonic scaler.
Afterwards, the teeth were placed for one hour in a 3\% hydrogen peroxide ultrasonic bath and then stored in 70\% ethanol {[}\protect\hyperlink{ref-1HkTk4VKM}{5}, \protect\hyperlink{ref-OttTicsv}{12}, \protect\hyperlink{ref-amSKhLb1}{15}, \protect\hyperlink{ref-ZTgDEakx}{17}, \protect\hyperlink{ref-1Gn1pVZXt}{19}{]}.

@@

\hypertarget{declarations}{%
\subsection{Declarations}\label{declarations}}

\hypertarget{ethics-approval-consent-to-participate-and-consent-for-publication}{%
\subsubsection{Ethics approval, consent to participate and consent for publication}\label{ethics-approval-consent-to-participate-and-consent-for-publication}}

The 104 human permanent mandibular canines in this study were collected from university medical centers in southwest Germany and Switzerland.
All included teeth were extracted for reasons unrelated to the study and are so-called excess material.
This makes an ethics approval unnecessary for the purpose of this investigation, as well as consent to participate and for publication.

\hypertarget{availability-of-data-and-materials}{%
\subsubsection{Availability of data and materials}\label{availability-of-data-and-materials}}

Two tooth datasets are available to download in the GitHub repository with the analysis notebook {[}\protect\hyperlink{ref-tZRGGuMm}{22}{]}.
The other datasets used for the current study are available from the corresponding author on reasonable request.

\hypertarget{competing-interests}{%
\subsubsection{Competing interests}\label{competing-interests}}

\hypertarget{funding}{%
\subsubsection{Funding}\label{funding}}

This research did not receive any specific grant from funding agencies in the public, commercial, or not-for-profit sectors.

\hypertarget{authors-contribution}{%
\subsubsection{Authors' contribution}\label{authors-contribution}}

@@

\hypertarget{acknowledgements}{%
\subsubsection{Acknowledgements}\label{acknowledgements}}

We thank Oleksiy-Zakhar Khoma (Institute of Anatomy, University of Bern, Switzerland) for designing \href{https://github.com/TomoGraphics/Hol3Drs/blob/master/STL/tooth_holder3.stl}{the sample holder that we used for scanning the teeth}.
We thank Jennifer Fazzari (Institute of Anatomy, University of Bern, Switzerland) for proof-reading the manuscript.
We thank Andrea Anderegg and Michael Stiebritz (both Department of Restorative, Preventive and Pediatric Dentistry, School of Dental Medicine, University of Bern, Switzerland) and Valentin Djonov (Institute of Anatomy, University of Bern, Switzerland) for their kind support.
